Carnegie SEM

### 2017-10-27 analysis/SEMwriteup.Rmd
some more changes to the writeup

```{r, echo=FALSE}
cc2015_matrix2 <- as.matrix(cc2015_r[-c(1:3)])
corrmatrix <- Hmisc::rcorr(cc2015_matrix2)
corrplot::corrplot(corrmatrix$r, order="hclust")
```

@@

```{r}
model2 <- '
Aggregate=~HUM_RSD+OTHER_RSD+SOCSC_RSD+STEM_RSD+PDNFRSTAFF+S.ER.D+NONS.ER.D
PerCap=~PDNRSTAFF_PC+S.ER.D_PC+NONS.ER.D_PC

@@

S.ER.D~~S.ER.D_PC
NONS.ER.D~~NONS.ER.D_PC
PDNFRSTAFF~~PDNRSTAFF_PC
PDNFRSTAFF~~S.ER.D
HUM_RSD~~SOCSC_RSD
S.ER.D~~STEM_RSD
PDNRSTAFF_PC~~S.ER.D_PC
NONS.ER.D~~HUM_RSD
NONS.ER.D~~SOCSC_RSD
'

@@

```{r}
###lets see if we can calculate distances with schools
Score_dist <- as.data.frame(cbind(CCScores_r_cov_scale, PC_pred_scale))
rownames(Score_dist) <- cc2015_new$NAME

@@

ggplot(Score_dist) + geom_point(aes(x = Aggregate, y = PerCap, color = CCScores_r_cov$rate_2015))

#calculate distances 
get_Distance <- function(x1,x2,y1,y2){

 #calculates the euclidean distance between the points  
  dist <- sqrt((x2 - x1)^2 + (y2 - y1)^2)
 #returns distances
  return(dist)
}

@@

for (j in 1:nrow(Score_dist)){
  Score_dist$Distance[j] <- get_Distance(x1 = Score_dist[j,"Aggregate"],x2 = Score_dist[j,"PC_ag_pred"],y1 = Score_dist[j,"PerCap"],y2 = Score_dist[j,"PC_percap_pred"])
}

@@

#show biggest movers on Carnegie Classifications Plot
ggplot(Score_dist) + geom_point(aes(x = PC_ag_pred, y = PC_percap_pred, color = as.factor(round(Distance)))) + ggtitle("Biggest Movers")

@@

>>>>

```{r, echo=FALSE}
par(mfrow = c(1,1))
cc2015_matrix2 <- as.matrix(cc2015_r[-c(1:3)])
corrmatrix <- Hmisc::rcorr(cc2015_matrix2)
corrplot::corrplot(corrmatrix$r, order="hclust")
```

@@

```{r}
model2 <- '
Aggregate=~HUM_RSD+OTHER_RSD+SOCSC_RSD+STEM_RSD+PDNFRSTAFF+S.ER.D+NONS.ER.D
PerCap=~PDNFRSTAFF_PC+S.ER.D_PC+NONS.ER.D_PC

@@

S.ER.D~~S.ER.D_PC
NONS.ER.D~~NONS.ER.D_PC
PDNFRSTAFF~~PDNFRSTAFF_PC
PDNFRSTAFF~~S.ER.D
HUM_RSD~~SOCSC_RSD
S.ER.D~~STEM_RSD
PDNFRSTAFF_PC~~S.ER.D_PC
NONS.ER.D~~HUM_RSD
NONS.ER.D~~SOCSC_RSD
'

@@

```{r}
#Generates PCA scores for comparison
PC_ag <- prcomp(cc2015.r[,-c(1,2,3,4)], scale = TRUE)
PC_percap <- prcomp(cc2015percap.r, scale = TRUE)

PC_ag_pred <- predict(PC_ag)[,1]
PC_percap_pred <- predict(PC_percap)[,1]

PC_pred_scale <- apply(as.data.frame(cbind(PC_ag_pred,PC_percap_pred)),2,scale)

#predicts the scores
CCScores_r_cov <- as.data.frame(lavaan::predict(lavaan_sem_r_cov))
CCScores_r_cov_scale <- apply(CCScores_r_cov[,c(1,2)], 2, scale)

###lets see if we can calculate distances with schools
Score_dist <- as.data.frame(cbind(CCScores_r_cov_scale, PC_pred_scale))
rownames(Score_dist) <- cc2015_new$NAME

@@

ggplot(Score_dist) + geom_point(aes(x = Aggregate, y = PerCap, color = CCScores_r_cov$rate_2015))
```

#Which Schools Move the Most?
The plot below shows which schools move the most relative to their location on the Carnegie classifications. These movements are not systematic; while most of the schools that move the most are near the center, they are not all occurning in one classification level.

```{r,distances, fig.align = 'center'}
#calculate distances 
get_Distance <- function(x1,x2,y1,y2){

 #calculates the euclidean distance between the points  
  dist <- sqrt((x2 - x1)^2 + (y2 - y1)^2)
 distx <- x2 - x1
 disty <- y2 - y1
  #returns distances
  return(list( dist = dist, distx = distx, disty = disty ))
}

@@

for (j in 1:nrow(Score_dist)){
  Score_dist$Distance[j] <- get_Distance(x1 = Score_dist[j,"Aggregate"],x2 = Score_dist[j,"PC_ag_pred"],y1 = Score_dist[j,"PerCap"],y2 = -Score_dist[j,"PC_percap_pred"])$dist

  }

@@

#show biggest movers on Carnegie Classifications Plot
ggplot(Score_dist) + geom_point(aes(x = PC_ag_pred, y = -PC_percap_pred, color = as.factor(floor(Distance)))) + ggtitle("Biggest Movers")

@@

#Additional Ways to Visualize These Models
There are two ways that we can visualize these plots. One uses GGplot to create path diagrams in a ggplot-based framework. These are similar to the SemPlot package but may allow for a bit more post processing of the plots. The unfortunate aspect of these plots is that they have to be put together piece by piece - there's no function that directly interfaces with lavaan to get the solution. 
```{r}
#add code to produce a path diagram here
```

An alternative way to look at these models would be to uses the Onyx package, which interfaces with the lavaan code but allows for you to make changes to the path diagrams in a separate window.  These are not great for this kind of context but they are useful for creating pretty-looking path diagrams. Onyx has to be installed from GitHub to work, but other than that it interfaces nicely with lavaan objects. 
```{r}
library(onyxR)
onyx(lavaan_sem_r_cov)
onyx(lavaan_sem_b)
onyx(lavaan_sem_r)
```
